---
title: "Kinship Tutorial"
author: "Morten Krebs"
date: "10/3/2022"
output: 
    html_document:
        theme: flatly
        toc: yes
vignette: >
    %\VignetteIndexEntry{PAFGRS Vignette Kinship Tutorial}
    %\VignetteEngine{knitr::rmarkdown}
    %\usepackage[utf8]{inputenc}
    %\SweaveUTF8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a short demonstration of how to compute the kinship matrix when predigrees are very large

# Load PAFGRS package

```{r, PAFGRS}
library(PAFGRS)

```

# Load family data
For demonstating this we load the publicly available data from the Minnesota Breast Cancer Family Study.
```{r, data}
data("minnbreast")

head(minnbreast[minnbreast$famid==4,1:7])
```



# Many small pedigrees 

For small pedigrees kinship can be efficiently estimated for using the kinship2 r-package:  

```{r simple}
data(minnbreast)
bpeds <- with(minnbreast,pedigree(id, fatherid, motherid, sex, affected=proband, famid=famid))
k <- kinship(bpeds)
dim(k)
```

# A single large pedigree
We can explore the idea of a single large pedigree by looking not looking at the famid:

```{r single-ped}
single_ped <- with(minnbreast[minnbreast$famid %in% 4:200,],pedigree(id, fatherid, motherid, sex, affected=proband))
single_ped
system.time(k <- kinship(single_ped))
dim(k)
format(object.size(k),units="Mb")
```

This object will quickly become very hard to store in memory. However we can take advantage of the fact that the matrix is very sparse: 

```{r single-ped-sparse}
system.time(k_sparse <- kinship_sparse(single_ped))
dim(k_sparse)
format(object.size(k_sparse),units="Mb")
```

Further, we can also estimate kinship coefficients for pedigrees at this scale by a path counting algorithm :

```{r single-ped-sparse-path}
system.time(k_sparse_path <- kinship_path(single_ped))

dim(k_sparse_path)
format(object.size(k_sparse_path),units="Mb")

identical(k_sparse_path,k_sparse)

```






