---
title: "Tutorial"
author: "Morten Krebs"
date: "9/28/2022"
output: 
    html_document:
        theme: flatly
        toc: yes
vignette: >
    %\VignetteIndexEntry{PAFGRS Vignette Tutorial}
    %\VignetteEngine{knitr::rmarkdown}
    %\usepackage[utf8]{inputenc}
    %\SweaveUTF8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load PAFGRS package

```{r, PAFGRS}
library("PAFGRS")
```

# Load family data
For a demonstation of the PAFGRS method we load the publicly available data from the Minnesota Breast Cancer Family Study.
```{r, data}
data("minnbreast")

head(minnbreast[minnbreast$famid==4,1:7])
```

# Compute a kinship matrix for one family
Using the \code{kinship2} package, we can compute the kinship matrix as follows: 

```{r, kinship}
ped_data <- minnbreast[minnbreast$famid==4,c(1:7,14)]
# generate pedigree
ped <- pedigree(id = ped_data$id,dadid = ped_data$fatherid,momid = ped_data$motherid,
                sex = ped_data$sex,affected = ped_data$cancer&ped_data$sex=="F", status=)
# compute kinsip matrix: 
k <- kinship2::kinship(ped)

plot(ped)
```


# Generate covariance matrix:
Under simplifying assumption (e.g. no shared environment, no inbreeding and no assortative mating) and a known heritability, the covariance in liabilities can be specified as: 
```{r, covmat}
r <- k*2
h2= 0.5 
covmat <- h2*r 
diag(covmat) <- 1
```

# Estimate liability of one individual:
## Naïve estimator 
A naiv estimator of the genetic liability of the first individual, assuming that all individuals has the same risk can be computed as:
```{r, fgrs-full-obs}
covmat <- covmat[c(3,(1:nrow(covmat))[-3]),c(3,(1:nrow(covmat))[-3])]
covmat[1,1] <- h2 
ped_data[ped_data$sex=="M",]$cancer <- 0 
status <- ped_data[-3,]$cancer
prev <- .05
thr <- qnorm(1-prev)

pa_fgrs(rel_status = status, covmat = covmat ,thr = thr)
```

## Sex-specific 
If we think that some individuals have a higher threshold than others, we can specify it as follows:
```{r, fgrs-sex}
ped_data$prev <- .1
ped_data$prev[ped_data$sex=="M"] <- 0
ped_data$thr <- qnorm(1-ped_data$prev)

pa_fgrs(rel_status = status, covmat = covmat,rel_thr = ped_data$thr[-3])
```

## Taking censoring into account 
Assuming that we know the cumulative incidence curve we can assign the mixture distribution for the censored individuals:

```{r, fgrs-age}
plot(sapply(1:100,function(x) {
  w=(x -20 )/80
  ifelse(w<0,0,w)*.1}),type = "l",ylab = "Cumulative Incidence in Females",xlab = "Age")
ped_data$w <-(ped_data$endage -20 )/80
ped_data$w[ped_data$w<0] <- 0
ped_data$w[ped_data$cancer==1] <- 1

pa_fgrs(rel_status = status, covmat = covmat,rel_thr = ped_data$thr[-3],rel_w = ped_data$w[-3])
```
